---
title: "Untitled"
author: "Ellie "hacker(wo)man supreme" Heffernan"
date: "6/7/2021"
output: html_document
---

#Pull data on number of people in each north carolina census tract who don't have health insurance. 
```{r}
nc_healthcare_65_older <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B27010_066E"),region="tract",regionin="state:37")
nc_healthcare_35_64 <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B27010_050E"),region="tract",regionin="state:37")
nc_healthcare_19_34 <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B27010_033E"),region="tract",regionin="state:37")
nc_healthcare_under_19 <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B27010_017E"),region="tract",regionin="state:37")




```


#Left_join those four dataframes and make a new column that is their sum, aka the total number of people in each census tract who don't have health insurance, regardless of age. 
```{r}
boo <- left_join(nc_healthcare_65_older,nc_healthcare_35_64)

boo2 <- left_join(boo,nc_healthcare_19_34)

boo3 <- left_join(boo2,nc_healthcare_under_19)

nc_total_without_healthcare <- boo3 %>% mutate(total_without_healthcare = B27010_066E+B27010_050E+B27010_033E+B27010_017E)
```





#Adjust this graph per capita.  
```{r}
#Make a table showing the total number of people in each census tract
nc_total <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B01003_001E"),region="tract",regionin="state:37")

#Join nc_total with nc_total_without_healthcare and create a new column dividing the total number of people in each census tract without healthcare (total_without_healthcare) by the total number of people in each census tract (B01003_001E)
nc_healthcare_adjusted<- left_join(nc_total_without_healthcare,nc_total) %>% mutate(healthcare_adjusted = total_without_healthcare/B01003_001E) 

#Now get rid of the original alphabet soup columns. 
nc_healthcare_adjusted <- nc_healthcare_adjusted %>% select(state, county, tract,healthcare_adjusted)

#View(nc_healthcare_adjusted)
```

#Join this with each of the shapefiles you made. This will connect that data solely to the census tracts in each county. 
```{r}
#Rename some columns, so joining is easier
#orange_census_tracts <- orange_census_tracts %>% rename(tract = TRACTCE)
#durham_census_tracts <-durham_census_tracts %>% rename(tract = TRACTCE)
#wake_census_tracts <- wake_census_tracts %>% rename(tract = TRACTCE)

#Filter only for each county's census tracts. 
orange_healthcare_adjusted <- nc_healthcare_adjusted %>% filter(county == "135")
wake_healthcare_adjusted <- nc_healthcare_adjusted %>% filter(county == "183")

#We came back here and changed two things to make it easier for readers to comprehend our interactive map: limiting numbers to 2 digits after the decimal point and converting from decimal to percent. 
durham_healthcare_adjusted <- nc_healthcare_adjusted %>% filter(county == "063") %>% mutate(healthcare_adjusted_percent = healthcare_adjusted*100)

  durham_healthcare_adjusted$healthcare_adjusted_percent <- round(durham_healthcare_adjusted$healthcare_adjusted_percent,digits = 2) 



#Making dataframes
orange_map4 <- inner_join(orange_census_tracts,orange_healthcare_adjusted)
durham_map4 <- inner_join(durham_census_tracts,durham_healthcare_adjusted)
wake_map4 <- inner_join(wake_census_tracts,wake_healthcare_adjusted)



```

#Creating static maps for each county 
```{r}
#Durham
ggplot(orange_map4)+geom_sf(aes(fill=healthcare_adjusted))+theme_pander()+theme(axis.text = element_blank(),axis.ticks = element_blank())+ theme(panel.grid.major=element_line(colour="transparent"))+scale_fill_distiller(palette="Blues",direction=1,name ="Residents without health insurance (%)") + labs(title = "Residents without health insurance in Orange County",caption = "Source: U.S. Census/ACS 2019")

#Orange
ggplot(durham_map4)+geom_sf(aes(fill=healthcare_adjusted))+theme_pander()+theme(axis.text = element_blank(),axis.ticks = element_blank())+theme(panel.grid.major=element_line(colour="transparent"))+scale_fill_distiller(palette="Blues",direction=1,name ="Residents without health insurance (%)") + labs(title = "Residents without health insurance in Durham County",caption = "Source: U.S. Census/ACS 2019")

#Wake
ggplot(wake_map4)+geom_sf(aes(fill=healthcare_adjusted))+theme_pander()+theme(axis.text = element_blank(),axis.ticks = element_blank())+   theme(panel.grid.major=element_line(colour="transparent"))+scale_fill_distiller(palette="Blues",direction=1,name ="Residents  without health insurance (%)") + labs(title = "Residents without health insurance in Wake County",caption = "Source: U.S. Census/ACS 2019")
```





#Plotting Durham on interactive map. 
```{r}

#Outlining the tracts 
durham_tracts <- tracts(state = "NC",county = "Durham",cb = T)

#Combining the dataframes
durham_map_interactive3 <- geo_join(durham_tracts,durham_healthcare_adjusted,"TRACTCE","tract") %>% filter(NAME!="9801")


#Creating a color palette 
pal <-colorNumeric(c("darkmagenta","yellow"),domain=durham_map_interactive3$healthcare_adjusted_percent)

#Creating a popup
popup_durham_healthcare <- paste0("Census Tract: ",as.character(durham_map_interactive3$NAME),"<br/> Residents without health insurance (%) :" ,as.character(durham_map_interactive3$healthcare_adjusted_percent))

#Creating the map 
durham_interactive_health_insurance_map <- leaflet() %>% addProviderTiles("CartoDB.Positron") %>% setView(-78.999,36.055,zoom =11) %>% addPolygons(data = durham_map_interactive3,fillColor =~pal(durham_map_interactive3$healthcare_adjusted_percent),fillOpacity = 1.5,weight = .2,smoothFactor=.2,popup =~ popup_durham_healthcare) %>% addLegend(pal=pal,values=durham_map_interactive3$healthcare_adjusted_percent,position = "bottomright",title = "<br/> Residents without <br/> health insurance  <br/> in Durham County (%)")




```

#Exporting interactives & tables that we need
```{r}
saveWidget(durham_interactive_health_insurance_map,"durham_interactive_health_insurance_map.html",selfcontained = TRUE)
write.csv(durham_healthcare_adjusted,"durham_healthcare_adjusted.csv")
```

































