---
title: '# of Black residents (Mapped)'
author: "Ellie "hacker(wo)man supreme" Heffernan"
date: "6/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


#Create a dataframe showing the total number of Black people in each census tract in the state.
```{r}
nc_race <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B02009_001E"),region="tract",regionin="state:37")
```


#Adjust this graph per capita. 
```{r}
#Make a table showing the total number of people per census tract in NC
nc_total <- getCensus(name = "acs/acs5",vintage = 2019,vars =c("B01003_001E"),region="tract",regionin="state:37")

#Join nc_total with nc_race and create a new column dividing the total number of Black people in each tract (B02009_001E) by the total number of people in each tract (B01003_001E). Black is defined as "Black alone or in combination with one or more races," since that is the definition the census api key used.  
nc_race_adjusted <- left_join(nc_total,nc_race) %>% mutate(black_adjusted = B02009_001E/B01003_001E)

#Now get rid of the original alphabet soup columns. 
nc_race_adjusted <- nc_race_adjusted %>% select(state, county, tract,black_adjusted)
 
#View(nc_race_adjusted)
```



#Join this with each of the shapefiles you made in the set-up file This will connect that data solely to the census tracts in those three counties. 
```{r}
#Rename some columns, so joining is easier
orange_census_tracts <- orange_census_tracts %>% rename(tract = TRACTCE)
durham_census_tracts <-durham_census_tracts %>% rename(tract = TRACTCE)
wake_census_tracts <- wake_census_tracts %>% rename(tract = TRACTCE)

#Filter only for each county's census tracts. 
orange_race_adjusted <- nc_race_adjusted %>% filter(county == "135")
durham_race_adjusted <- nc_race_adjusted %>% filter(county == "063") %>% mutate(black_adjusted_percent = black_adjusted*100)
  durham_race_adjusted$black_adjusted_percent <- round(durham_race_adjusted$black_adjusted_percent,digits = 2) 
  #I came back and made some edits here to pretty up the interactive map I made later. Changing from decimal to percent and only allowing two digits after the decimal 
wake_race_adjusted <- nc_race_adjusted %>% filter(county == "183")


#Joining
orange_map1 <- inner_join(orange_census_tracts,orange_race_adjusted)
durham_map1 <- inner_join(durham_census_tracts,durham_race_adjusted)
wake_map1 <- inner_join(wake_census_tracts,wake_race_adjusted)



```

#Plotting data for each Triangle county on static maps.
```{r}


#Orange
ggplot(orange_map1,)+geom_sf(aes(fill=black_adjusted))+theme_pander()+ theme(panel.grid.major=element_line(colour="transparent"))+ theme(axis.text = element_blank(),axis.ticks = element_blank())+                                                                   scale_fill_distiller(palette="Blues",direction=1,name ="% of residents that are Black") + labs(title = "Black residents in Orange County census tracts ",caption = "Source: U.S. Census/ACS 2019")

#Durham
ggplot(durham_map1)+geom_sf(aes(fill=black_adjusted))+theme_pander()+ theme(axis.text = element_blank(),axis.ticks = element_blank())+ theme(panel.grid.major=element_line(colour="transparent"))+scale_fill_distiller(palette="Blues",direction=1,name ="% of residents that are Black") + labs(title = "Black residents in Durham County census tracts",caption = "Source: U.S. Census/ACS 2019")

#Wake
ggplot(wake_map1)+geom_sf(aes(fill=black_adjusted))+theme_pander()+ theme(axis.text = element_blank(), axis.ticks = element_blank())+     theme(panel.grid.major=element_line(colour="transparent"))+scale_fill_distiller(palette="Blues",direction=1,name ="% of residents that are Black") + labs(title = "Black residents in Wake County census tracts",caption = "Source: U.S. Census/ACS 2019")


```
#Note to self: Grey tracts have NO residents. In Durham, this tract is mainly RTP & the Durham freeway. In Wake, the tracts are the Airport & William B. Umstead State Park. 


#Plotting Durham on interactive map. 
```{r}

#Outlining the tracts 
durham_tracts <- tracts(state = "NC",county = "Durham",cb = T)

#Combining the dataframes
durham_map_interactive1 <- geo_join(durham_tracts,durham_race_adjusted,"TRACTCE","tract") %>% filter(NAME!="9801")

#Creating a color palette 
pal <-colorNumeric(c("darkmagenta","yellow"),domain=durham_map_interactive1$black_adjusted_percent)

#Creating a popup
popup_durham_black_pop <- paste0("Census Tract: ",as.character(durham_map_interactive1$NAME),"<br/> Residents that are Black (%): " ,as.character(durham_map_interactive1$black_adjusted_percent))

#Making the map 
durham_interactive_race_map <- leaflet() %>% addProviderTiles("CartoDB.Positron") %>% setView(-78.999,36.055,zoom =11) %>% addPolygons(data = durham_map_interactive1,fillColor =~pal(durham_map_interactive1$black_adjusted_percent),fillOpacity = 1.5,weight = .2,smoothFactor=.2,popup =~ popup_durham_black_pop) %>% addLegend(pal=pal,values=durham_map_interactive1$black_adjusted_percent,position = "bottomright",title = "<br/> Distribution of <br/> Black residents <br/> in Durham County (%)")


 
```


#Exporting your interactive map
```{r}
library(htmlwidgets)
saveWidget(durham_interactive_race_map,"durham_interactive_race_map.html",selfcontained = TRUE)
```











