---
title: "Heat Maps"
author: "Ellie "hacker(wo)man supreme" Heffernan"
date: "6/3/2021"
output: html_document
---

#Load packages & make sure your census key is loaded -- (I don't entirely know what all of this means. I just copy & paste the set-up that my professor recommended)
```{r}
#Load packages
library(tigris)
library(pacman)
library(censusapi)
library(ggthemes)
library(tidyverse)
library(stringr)
p_load(leaflet,glue,dplyr,sf,tmap,tmaptools,tidycensus,ggmap,htmltools, htmlwidgets)
p_load_gh(c("walkerke/tigris","bhaskarvk/leaflet.extras"))
options(tigris_use_cache = TRUE)


#Load census key
census_key <- "ee879694947d7f51fb7006bed7d04d8d980a42fc"
Sys.setenv(CENSUS_KEY=census_key)
install = TRUE
```


#Make shapefiles for Durham, Wake, and Orange County. This is just the outlines. 
```{r}
durham_census_tracts <-tracts("NC", county = "Durham", year = 2019,class = "sf")
                          
wake_census_tracts <-tracts("NC", county = "Wake", year = 2019,class = "sf")  

orange_census_tracts <- tracts ("NC",county = "Orange",year = 2019, class = "sf")


```


#KEEP THESE TWO VARIABLES! "apis" shows you the different surveys you can pull from (i.e. ACS, Decennial Census) & acs_vars shows you the many different variables you can pull from the American Community Survey (2019)
```{r}
#apis <-listCensusApis()
#acs_vars <-listCensusMetadata(name ="acs/acs5",type ="variables",vintage = "2019")

write.csv(acs_vars,"acs_var.csv")
write.csv(apis,"apis.csv")

```

#List of variables and vintages used
```{r}
#These all came from the American Community Survey 2019  (acs/acs5,  vintage = "2019")

#Variables Used: 


#B02009_001E: Total number BLACK OR AFRICAN AMERICAN ALONE OR IN COMBINATION WITH ONE OR MORE OTHER RACES
#For adjustment: B01003_001E: Total population 



#B25035_001E: Median year structure built 


#B27010_066E: Total number 65 and older without health insurance
#B27010_050E: Total number 35 to 64 without health insurance
#B27010_033E: Total number 19 to 34 without health insurance
#B27010_017E: Total number under 19 without health insurance
#For adjustment: 01003_001E: Total population 


```




